---
title: "Housing Bubble Shiny App"
author: "Sebastian Sanin"
date: "20/10/2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Objective
This presentation shows the use of my new Shiny app: *Housing Bubbles and Crises periods around the Globe*

<div align="center">
<img src=fig/app1.png width=600 height=300>
</div>

This app shows the housing bubble periods for different countries and plots the price indicators with bubbles as shaded areas.

The app is completely functional. When bubbles are not identified nor table nor plot is depicted and an error appears. Change your country or indicator selection in this case.

## Usage of the App
In order to compute housing bubble periods you will have to choose the following two variables:

- **Country**: From the list you need to select the country code that you want to work with.
- **Indicator**: From the list you have two choose between Real house price index (rhpi) and Real house price to disposable income index (rhpi_rpdi).

Finally you have to click on the **Plot Bubbles** button and wait some seconds.

## Example
Suppose you select **Country:** *WRD* that refers to world and **Indicator:** *rhpi_rpdi* that refers to real house price to disposable income index. You will obtain:

```{r, echo=F}
suppressPackageStartupMessages(suppressWarnings(library(scales)))
suppressPackageStartupMessages(suppressWarnings(library(plyr)))
suppressPackageStartupMessages(suppressWarnings(library(dplyr)))
suppressPackageStartupMessages(suppressWarnings(library(data.table)))
suppressPackageStartupMessages(suppressWarnings(library(ggplot2)))
suppressPackageStartupMessages(suppressWarnings(library(plotly)))
suppressPackageStartupMessages(suppressWarnings(library(psymonitor)))

ctrycodes <- data.table(read.csv("country_codes.csv",sep=";"))
df <- data.table(read.csv("bubble_file_new.csv",sep=";",
                          colClasses = c("POSIXct",rep("factor",5),"numeric")))
df$date <- as.Date(df$date)

indicator <- "rhpi_rpdi"
ctry <- "WRD"

df2 <- df[variable==indicator&country==ctry,]

y        <- df2$value
obs      <- length(y)
swindow0 <- floor(obs*(0.01 + 1.8/sqrt(obs)))
dim      <- obs - swindow0 + 1
Tb       <- 24 + swindow0 - 1
# Estimate PSY statistics and CVs
bsadf          <- PSY(y, swindow0)
quantilesBsadf <- cvPSYwmboot(y, swindow0, Tb=Tb, nboot = 50, nCores = 2)
quantile95     <- quantilesBsadf %*% matrix(1, nrow = 1, ncol = dim)
# locate bubble/crisis dates
ind95        <- (bsadf > t(quantile95[3, ])) * 1
monitorDates <- df2$date[swindow0:obs]
OT           <- locate(ind95, monitorDates)
# Show bubble/crisis periods
datebub <- disp(OT, obs)
t(datebub)
```


```{r, echo=F, fig.width = 4,fig.height=3}
# Graphics
ggplot(df2)+geom_line(aes(x=as.Date(date),y=value))+
  geom_rect(data=datebub,aes(xmin=as.Date(start),xmax=as.Date(end)),ymin=-Inf,ymax=Inf,alpha=0.3,fill="red")+
  scale_x_date(labels=date_format ("%Y-%m"),breaks=date_breaks("4 year"))+  
  scale_y_continuous(limits = c(min(df2$value)-20,max(df2$value)+20))+
  theme_bw()+labs(title=paste("Housing bubbles or crises in: ",ctrycodes[country_code==ctry,country_name],sep=""),
                  x="",y=ifelse(indicator=="rhpi_rpdi","Real house price to income index","Real house price index"))+
  theme(axis.text.x = element_text(angle = 90))

```

## Results
The resulting Bubble/Crisis Periods table shows the *start* and *end* of these events.
<div align="center">
<img src=fig/app2.png width=600 height=300>
</div>
In the lower part you will see a plot of the selected indicator and the bubble period in a light-red shaded area.
<a href='https://ssaninre22.shinyapps.io/HousePriceBubbleApp/' title=''>My app</a>

**THANKS FOR YOUR ATTENTION AND ENJOY MY APP**





